---
title: "800-171 Analysis"
output: 
  pdf_document: 
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(readr)
library(tidyverse)
library(ggplot2)
library(igraph)
library(dplyr)
```     
 

#Basic Analysis 
  
##Data Loading

Loading the SP-800-171 excel file into R for analysis. 
```{r}
#Julia Smadja's working directory, what's in quotation marks must changed
Security_Requirements <- read_excel("800-171 Appendix D.xlsx") 
head(Security_Requirements)
```   
    
        
##Data Cleaning
   
   
```{r}
Security_Requirements <- Security_Requirements %>%
  #renaming variables so there is no space
  select(`SR_Family` = `SR Family`, 
         `SR_Family_Name` = `SR Family Name`, 
         `Type_of_SR` = `Type of SR`, 
         `SP_800-171` = `SP 800-171`,
         `SP_800-171_SR` = `SP 800-171 SR`,
         `800-53_Family` = `800-53 Family`,
         `NIST_SP_800-53` = `NIST SP 800-53`, 
         `NIST_SP_800-53_Relevant_Security_Controls` = `NIST SP 800-53 Relevant Security Controls`,
         `ISO/IEC_27001_Relevant_Security_Controls` = `ISO/IEC 27001 Relevant Security Controls`) 
  
  
`SP_800-171 Cleaned` <- Security_Requirements %>%
  select(SR_Family, SR_Family_Name, `SP_800-171`, `800-53_Family`, `NIST_SP_800-53`) %>%
  distinct()

head(`SP_800-171 Cleaned`)
#View(`SP_800-171 Cleaned`)
```
     
     
#Data Analysis    
  
The purpose of this analysis is to understand the different security requirements (SP 800-171) and their families and their relationships with the security controls in NIST Special Publication 800-53. An additional visual analysis is done using Tableau for ease of interpretation. 

**Number of Security Requirement Families for SP-171 and their names?**    
    
```{r}
Count_SRFamily <- `SP_800-171 Cleaned` %>%
  summarise(count_SRFamily = n_distinct(SR_Family))

head(Count_SRFamily)

Names_SRFamily <- `SP_800-171 Cleaned` %>%
  group_by(SR_Family, SR_Family_Name) %>%
  summarise()

head(Names_SRFamily)
```
    
There are 14 security requirement families to protect the confidentiality of CUI in nonfederal systems and organizations.  

As the analysis moves forward, the analysis dives deeper into the security requirements, how many there are and how they are related to their families.
   
**Number of Security Requirement for SP-171 and their "names"?**    
    
```{r}
Count_SR <- `SP_800-171 Cleaned` %>%
  summarise(count_SR = n_distinct(`SP_800-171`))

head(Count_SR)

Names_SR <- `SP_800-171 Cleaned` %>%
  group_by(`SP_800-171`) %>%
  summarise()

head(Names_SR)
```   
    
There is a total of 110 security requirements within those families. 
   
**Security Requirement families with the number of security requirements (SP 800-171)?**   
     
```{r}
Count_SRFamily_SR <- `SP_800-171 Cleaned` %>%
  group_by(SR_Family) %>%
  summarise(count_of_SP_800_171 = n_distinct(SR_Family, `SP_800-171`)) %>%
  arrange(desc(count_of_SP_800_171 ))

head(Count_SRFamily_SR)  
   
Names_SRFamily_SR <- `SP_800-171 Cleaned` %>%
  group_by(SR_Family) %>%
  summarise(names_of_SP_800_171 = toString(`SP_800-171`)) %>%
  ungroup()

head(Names_SRFamily_SR)
```    
   
From the 110 security requirements, the analysis looks into how they are grouped within each security family. The security requirement family with the largest amount of security requirements is 3.01 (ACCESS CONTROL) with 22, followed by 3.13 (SYSTEM AND COMMUNICATIONS PROTECTION) with 16 and 3.05 (IDENTIFICATION AND AUTHENTICATION) with 11.    
   
The next steps in the analysis is to add the NIST 800-53 and understand the mapping of the security requirements to the relevant security controls (NIST 800-53).
   
      
**Security Requirement Families (SP 800-171) with the number of security control families (SP 800-53) and the names of those security control families.**    
      
```{r}
Count_SRFamily_SCFamily <- `SP_800-171 Cleaned` %>%
  group_by(SR_Family) %>%
  summarise(count_of_SP_800_53Families = n_distinct(`800-53_Family`)) %>%
  arrange(desc(count_of_SP_800_53Families))

Count_SRFamily_SCFamily

Names_SRFamily_SCFamily <- `SP_800-171 Cleaned` %>%
  group_by(SR_Family) %>%
  summarise(names_of_SP_800_53Families = toString(unique(`800-53_Family`))) %>%
  ungroup()

head(Names_SRFamily_SCFamily)
```      
   
Within the security requirement families there are only three families that have 2 security control families, 3.08, 3.12 and 3.13. All the other families only have one security control family. 3.08 has CP and MP, 3.12 has CA and PL, and 3.13 has SA and SC. 
    
**Security Requirement Families (SP 800-171) with the number of security controls (SP 800-53) and the names of those security controls.**    
      
```{r}
Count_SRFamily_SC <- `SP_800-171 Cleaned` %>%
  group_by(SR_Family) %>%
  summarise(count_of_SP_800_53 = n_distinct(`NIST_SP_800-53`)) %>%
  arrange(desc(count_of_SP_800_53))

Count_SRFamily_SC

Names_SRFamily_SC <- `SP_800-171 Cleaned` %>%
  group_by(SR_Family) %>%
  summarise(names_of_SP_800_53 = toString(unique(`NIST_SP_800-53`))) %>%
  ungroup()

head(Names_SRFamily_SC)
```      
   
Within the Security Requirement families, 3.01 has the largest amount of security controls with 28, followed by 3.13 with 16, 3.03 with 14, 3.04 with 13 and 3.05 with 11. 
     
**Security Requirement (SP 800-171) with the number of security control families (SP 800-53) and the names of those security control families.**   
       
```{r}
Count_SR_SCFamily <- `SP_800-171 Cleaned` %>%
  group_by(`SP_800-171`) %>%
  summarise(count_of_SP_800_53Families = n_distinct(`SP_800-171`, `800-53_Family`)) %>%
  arrange(desc(count_of_SP_800_53Families))

head(Count_SR_SCFamily)

Names_SR_SCFamily <- `SP_800-171 Cleaned` %>%
  group_by(`SP_800-171`) %>%
  summarise(names_of_SP_800_53Families = toString(unique(`800-53_Family`))) %>%
  ungroup()

head(Names_SR_SCFamily)
```        
       
There are 6 security requirements that have 2 different security controls, 3.12.1, 3.12.2, 3.12.3, 3.12.4, 3.13.1 and 3.13.2. 
           
           
**Security Requirement (SP 800-171) with the number of security controls (SP 800-53) and the names of those security controls?**   
   
```{r}
Count_SR_SC <- `SP_800-171 Cleaned` %>%
  group_by(`SP_800-171`) %>%
  summarise(count_of_SP_800_53 = n_distinct(`SP_800-171`,`NIST_SP_800-53`)) %>%
  arrange(desc(count_of_SP_800_53))

head(Count_SR_SC)

Names_SR_SC <- `SP_800-171 Cleaned` %>%
  group_by(`SP_800-171`) %>%
  summarise(names_of_SP_800_53 = toString(`NIST_SP_800-53`)) %>%
  ungroup()

head(Names_SR_SC)
```     
   
The security requirements with the most number of security controls are 3.3.1 with 6 and 3.3.2 with 6, followd by 3.6.1 with 5 and 3.6.2 with 5.   
   
##Exporting Data 
    
```{r}
#list_of_datasets <- list("SP_800-171 Original" = `Security_Requirements`,
#                         "SP_800-171 Cleaned" = `SP_800-171 Cleaned`, 
#                         "Names_SRFamily" = `Names_SRFamily`,
#                         "Names_SR" = `Names_SR`,
#                         "Count_SRFamily_SR" = `Count_SRFamily_SR`,
#                         "Names_SRFamily_SR" = `Names_SRFamily_SR`,
#                         "Count_SRFamily_SCFamily" = `Count_SRFamily_SCFamily`,
#                         "Names_SRFamily_SCFamily" = `Names_SRFamily_SCFamily`, 
#                         "Count_SRFamily_SC" = `Count_SRFamily_SC`, 
#                         "Names_SRFamily_SC" = `Names_SRFamily_SC`, 
#                         "Count_SR_SCFamily" = `Count_SR_SCFamily`, 
#                         "Names_SR_SCFamily" = `Names_SR_SCFamily`, 
#                         "Count_SR_SC" = `Count_SR_SC`, 
#                         "Names_SR_SC" = `Names_SR_SC`)

#library(xlsx)
#write.xlsx(list_of_datasets, "SP 800-171 Analysis.xlsx")

write_csv(`SP_800-171 Cleaned`, "SP_800-171 Cleaned.csv")
write_csv(`Names_SRFamily`, "Names_SRFamily.csv")
write_csv(`Names_SR`, "Names_SR.csv")
write_csv(`Count_SRFamily_SR`, "Count_SRFamily_SR.csv")
write_csv(`Names_SRFamily_SR`, "Names_SRFamily_SR.csv")
write_csv(`Count_SRFamily_SCFamily`, "Count_SRFamily_SCFamily.csv")
write_csv(`Names_SRFamily_SCFamily`, "Names_SRFamily_SCFamily.csv")
write_csv(`Count_SRFamily_SC`, "Count_SRFamily_SC.csv")
write_csv(`Names_SRFamily_SC`, "Names_SRFamily_SC.csv")
write_csv(`Count_SR_SCFamily`, "Count_SR_SCFamily.csv")
write_csv(`Names_SR_SCFamily`, "Names_SR_SCFamily.csv")
write_csv(`Count_SR_SC`, "Count_SR_SC.csv")
write_csv(`Names_SR_SC`, "Names_SR_SC.csv")
```

After exporting these excel files, combine these 13 different excel workbooks into the same workbook, SP 800-171 Analysis, with 13 different worksheets.    
Combine the Count_SR_SC and Names_SR_SC into the same sheet and Count_SRFamily_SR and Names_SRFamily_SR. 
Add the original SP 800-171 data set to the workbook, SP 800-171 Analysis.   
   
A visual analysis is done with Tableau. 
    
    
#Advanced Data Analysis 

Once the basic overall analysis is over a deeper analysis looking at the relationship with the SP 800-53 and the percentage of them being used within the security requirements.   
   
Loading in the 800-53 Security Control Data Set    
```{r}
Security_Controls <- read_excel("~/Desktop/GitHub/FedRAMP_800-53_Controls/800-53 subcontrol relations.xlsx")
head(Security_Controls)
```
   
      
##Data Cleaning  

```{r}
`SP_800-53 Cleaned` <- Security_Controls %>%
  filter(High == "Y") %>% #filtering to High Baseline 
  #renaming variables so there is no space and selecting the variables wanted
  select(`SC_Family` = `Family`, 
         `SC_Family_Name` = `Family Name`, 
         `Control_Name` = `Control Name`, 
         `Main_Control` = `Main Control`,
         `Control_Enhancement_Name` = `Control Enhancement Name`,
         `Control_Enhancement` = `Control Enhancement`,
         `Combined` = `Combined`) %>%
  distinct()

head(`SP_800-53 Cleaned`)
#View(`SP_800-53 Cleaned`)
```    
  
The data is manipulated in a way to best represent the relationship between the security controls and the security requirements.    

##Data Analysis   
    
    
**Number of Security Control Families for SP 800-53 and their names?**    
    
```{r}
Count_SCFamily <- `SP_800-53 Cleaned` %>%
  summarise(count_SCFamily = n_distinct(SC_Family))

head(Count_SCFamily)

Names_SCFamily <- `SP_800-53 Cleaned` %>%
  group_by(SC_Family, SC_Family_Name) %>%
  summarise()

head(Names_SCFamily)
```   
   
There are 17 Security Control families for SP 800-53.    
   
**Number of Security Controls for SP-800-53 (Combined) and their "names"?**    
    
```{r}
Count_SC <- `SP_800-53 Cleaned` %>%
  summarise(count_SC = n_distinct(`Combined`))

head(Count_SC)

Names_SC <- `SP_800-53 Cleaned` %>%
  group_by(`Combined`) %>%
  summarise()

head(Names_SC)
```   
   
There are 415 Security Controls within the 17 Security Control families. These are for FedRAMP High level.    
   
**Security Control families with the number of security controls (SP 800-53)?**   
     
```{r}
Count_SCFamily_SC <- `SP_800-53 Cleaned` %>%
  group_by(SC_Family) %>%
  summarise(count_of_SP_800_53 = n_distinct(SC_Family, `Combined`)) %>%
  arrange(desc(count_of_SP_800_53 ))

head(Count_SCFamily_SC)  
   
Names_SCFamily_SC <- `SP_800-53 Cleaned` %>%
  group_by(SC_Family) %>%
  summarise(names_of_SP_800_53 = toString(`Combined`)) %>%
  ungroup()

head(Names_SCFamily_SC)
```       
   
The Security Control family with the most Security Controls is AC with 54, followed by SI with 39, SC with 38, CM with 36 and CP with 35.   
    
      
##Exporting Data 
    
```{r}
write_csv(`SP_800-53 Cleaned`, "SP_800-53 Cleaned.csv")
write_csv(`Names_SCFamily`, "Names_SCFamily.csv")
write_csv(`Names_SC`, "Names_SC.csv")
write_csv(`Count_SCFamily_SC`, "Count_SCFamily_SC.csv")
write_csv(`Names_SCFamily_SC`, "Names_SCFamily_SC.csv")
``` 

Once exported, the different data sets are combined into one dataset with multiple different sheets.     
SC_Family_Name is turned into capitalized letters and the a IF statement, ISERROR and VLOOKUP is created to find the Security Controls that appear in the SP 800-171 and the ones that do not.    
Visuals are created with Tableau for analysis.    
    
#800-171 and 800-53 Comparison  
  
```{r}
`SP_800_171_Analysis` <- read_excel("SP 800-171 Analysis.xlsx")
`SP_800_53_Analysis_for_SP_800_171` <- read_excel("SP 800-53 Analysis for SP 800-171.xlsx")

head(SP_800_171_Analysis)
head(SP_800_53_Analysis_for_SP_800_171)
```

For this analysis, we would want to find the percentage of Security Controls that are within the SP-800-171 and analyze the percentage difference between the two.    
      
**Number of Security Control Families in SP 800-53 and the SP 800-171.**
```{r}
Count_SCFamily53 <- SP_800_53_Analysis_for_SP_800_171 %>%
  mutate(total_SCFamily53 = length(unique(SC_Family))) %>%
  select(SC_Family, total_SCFamily53) %>%
  distinct()

Count_SCFamily53

Count_SCFamily171 <- SP_800_171_Analysis %>%
  mutate(total_SCFamily171 = length(unique(`800-53 Family`))) %>%
  select(`800-53 Family`, total_SCFamily171) %>%
  distinct() %>%
  rename(`SC_Family` = `800-53 Family`)

Count_SCFamily171

Percentage_SCFamilies <- Count_SCFamily53 %>%
  left_join(Count_SCFamily171) %>% 
  mutate(percentage_SCFamilies = total_SCFamily171/total_SCFamily53) %>%
  select(-SC_Family) %>%
  distinct()
```
Both, SP 800-53 and SP 800-171 have the same number of Security Control families. This means that SP 800-171 uses the total amount (100%) of possible Security Control family possible.     

  
**Number of Security Control in SP 800-53 and the SP 800-171.**
```{r}
SP_800_53_Analysis_for_SP_800_171
Count_SC53 <- SP_800_53_Analysis_for_SP_800_171 %>%
  mutate(total_SC53 = length(unique(Combined))) %>%
  select(Combined, total_SC53) %>%
  rename(`NIST SP 800-53` = `Combined`) %>%
  distinct()

Count_SC53

SP_800_171_Analysis
Count_SC171 <- SP_800_171_Analysis %>%
  mutate(total_SC171 = length(unique(`NIST SP 800-53`))) %>%
  select(`NIST SP 800-53`, total_SC171) %>%
  distinct()

Count_SC171

Percentage_SC <- Count_SC53 %>%
  left_join(Count_SC171) %>% 
  mutate(percentage_SC = total_SC171/total_SC53) %>%
  select(-`NIST SP 800-53`) %>%
  distinct() %>%
  drop_na()
```    
SP 800-53 has a total of 415 Security Controls, while SP 800-171 has 127 of those 415 Security Controls. This means that 30.6% of the Security Controls are in SP 800-171.     
   
    
**Number of Security Control withing each family in SP 800-53 and the SP 800-171.**
```{r}
Count_SCFamily_SC53 <- SP_800_53_Analysis_for_SP_800_171 %>%
  group_by(SC_Family) %>%
  mutate(total_SCFamily_SC53 = length(SC_Family)) %>%
  select(SC_Family, total_SCFamily_SC53) %>%
  distinct()

Count_SCFamily_SC53

SP_800_171_Analysis
Count_SCFamily_SC171 <- SP_800_171_Analysis %>%
  group_by(`SR Family`, `800-53 Family`) %>%
  mutate(total_SCFamily_SC171 = length(unique(`NIST SP 800-53`))) %>%
  ungroup() %>%
  group_by(`SR Family`) %>%
  mutate(total_SRFamily_SC171 = length(unique(`NIST SP 800-53`))) %>%
  select(`800-53 Family`, total_SRFamily_SC171, total_SCFamily_SC171) %>%
  distinct() %>%
  rename(`SC_Family` = `800-53 Family`)

Count_SCFamily_SC171

Percentage_SCFamily_SC <- Count_SCFamily_SC53 %>%
  left_join(Count_SCFamily_SC171, by = "SC_Family") %>% 
  mutate(percentage_SCFamily = total_SCFamily_SC171/total_SCFamily_SC53) %>% 
         #percentage_SRFamily = total_SRFamily_SC171/total_SCFamily_SC53) %>%
  select(`SR Family`, SC_Family, total_SCFamily_SC53, total_SCFamily_SC171, percentage_SCFamily)
#total_SRFamily_SC171, percentage_SRFamily,

Percentage_SCFamily_SC 
```

**Number of Security Control withing each family in SP 800-53 and the SP 800-171 by Security Requirement Family.**
```{r}
#Count_SCFamily_SC53 <- SP_800_53_Analysis_for_SP_800_171 %>%
#  group_by(SC_Family) %>%
#  mutate(total_SCFamily_SC53 = length(SC_Family)) %>%
#  select(SC_Family, total_SCFamily_SC53) %>%
#  distinct()

#Count_SCFamily_SC53

#SP_800_171_Analysis
#Count_SCFamily_SC171_SRFamily <- SP_800_171_Analysis %>%
#  group_by(`SR Family`) %>%
#  mutate(total_SCFamily_SC171_SRFamily = length(unique(`NIST SP 800-53`))) %>%
#  select(`800-53 Family`, `SR Family`, total_SCFamily_SC171_SRFamily) %>%
#  distinct()

#Count_SCFamily_SC171_SRFamily

#Percentage_SCFamily_SC_SRFamily <- Count_SCFamily_SC171_SRFamily %>%
#  left_join(Count_SCFamily_SC53) %>% 
#  mutate(percentage_SC = total_SCFamily_SC171_SRFamily/total_SCFamily_SC53)

#Percentage_SCFamily_SC_SRFamily 
```   
    
   
##Exporting Data 
    
```{r}
write_csv(`Percentage_SCFamilies`, "Percentage_SCFamilies.csv")
write_csv(`Percentage_SC`, "Percentage_SC.csv")
write_csv(`Percentage_SCFamily_SC`, "Percentage_SCFamily_SC.csv")
``` 